### 2009-04-04 talk/icooolps2009-dotnet/paper.tex
some adjustments in the abstract

\begin{abstract}
The CLI, i.e. the virtual machine at the core of the .NET environment, is a
statically typed virtual machine.  Although it is being used as a target for
many different languages, it is hard to provide higly efficient implementation
of languages whose model is too different than the one of the VM.  In
particular, dynamically typed languages are typically order of magnitude
slower than, e.g., C\#.

\anto{XXX: should we cite the DLR? How to prove that DLR programs are much
  slower than C\#?}

Recent developments show that JIT compilers can exploit runtime type
information to generate quite efficient code.  Unfortunately, writing a JIT
compiler is far from being simple.  In this paper we report our positive
experience with automatic generation of JIT compilers as supported by the PyPy
infrastructure, by focusing on JIT compilation for .NET.

JIT compilation for .NET means to generate .NET bytecode at runtime, which
will further compiled down to machine code by .NET's own JIT compiler.  The
main and novel contribution of this paper is to show that this
\emph{two-layers JIT} technique is effective, and can give very high speedups,
up to make dynamic languages faster than C\# itself under certain
circumstances.

>>>>

\begin{abstract}
The Common Language Infrastructure (CLI) is a virtual machine expressly
designed for implementing statically typed languages as C\#, therefore
programs written in dynamically typed languages are typically much slower than C\# when executed on .NET.

\anto{XXX: should we cite the DLR? How to prove that DLR programs are much
  slower than C\#?}
\davide{We can try writing something in the conclusion}

Recent developments show that JIT compilers can exploit runtime type
information to generate quite efficient code.  Unfortunately, writing a JIT
compiler is far from being simple.

In this paper we report our positive
experience with automatic generation of JIT compilers as supported by the PyPy
infrastructure, by focusing on JIT compilation for .NET.
Following this approach, we have in fact added a second layer of JIT compilation, by allowing dynamic generation of more efficient .NET bytecode, which
in turn can be compiled to machine code by the .NET JIT compiler.

The main and novel contribution of this paper is to show that this
\emph{two-layers JIT} technique is effective, since programs written in dynamic languages 
can run on .NET as fast as (and in some cases even faster than) the equivalent C\# programs.
